Viscous Erosion in Porous Media

### 2018-03-04 docs/methods/erosion_methods.tex
NM

For numerical stability, we modify the erosion law \eqref{eqn:erosionModel} to include a term that penalizes curvature $\kappa$. The idea is to prevent regions of extremely high curvature from developing, as these regions would compromise the accuracy and stability of our fluid solver. We thus modify \eqref{eqn:erosionModel} follows
\begin{align}
\label{eqn:Vn}
\Vn = \atau + \eps \mean{\atau} \left( \frac{L }{2\pi} \kappa - 1 \right).
\end{align}
The parameter $\eps \ll 1$ controls the strength of the curvature-penalization. This term causes regions of high curvature to recede more rapidly and thus smooths out any high-frequency oscillations that may exist. Above, $\mean{\cdot}$ indicates a boundary mean. Since a closed body has mean curvature $\mean{\kappa} = 2\pi / L$, the penalty term has mean zero and and thus preserves area. The only source of material loss is therefore the shear-stress itself. In addition, we have scaled the penalty term on the mean absolute shear, $\mean{\atau}$, so that the ratio of smoothing-to-erosion always remains fixed, even as bodies vanish.

@@

The boundary of each body will be represented by a grid of collocation points. If we were to simply evolve the interfaces under \eqref{eqn:Vn2}, this grid would become distorted, which, if unchecked, could lead to severe tangling and numerical instability. We therefore introduce an artificial tangential velocity $\Vs$ that always keeps the grid points equally spaced in arclength \cite{hou-low-she1994}. The appropriate tangential velocity satisfies 
\begin{align}
\pderiv{\Vs}{\alpha} = \thalpha \Vn - \mean{\thalpha \Vn} \, ,
\end{align}
The evolution of the interface is then given by the combination of the normal and tangential velocity
\begin{align}
\dot{\xx}(t) = \Vn \nn + \Vs \ss \, .
\end{align}
Since $\Vs$ is tangential, it does not modify the physics of erosion, only the grid spacing.

@@

With the shear stress computed, we are now ready to numerically evolve the boundaries using the {\thL} framework. The main equations to be solved are \eqref{eqn:Levo1}--\eqref{eq:xsurf}.  In Section~\ref{sec:thetaL} we introduced a curvature-penalization term into the erosion law \eqref{eqn:Vn}. This term smooths out high-frequency oscillations, but also introduces {\em stiffness} into evolution equation \eqref{eqn:thetaevo1} for $\theta(\alpha,t)$. For numerical stability, we will treat the stiff term with an exponential integrator, while handling all other terms in \eqref{eqn:Levo1}--\eqref{eq:xsurf} with an explicit Runge-Kutta method. Ultimately, we will arrive at a method that is second-order in time.

>>>>

For numerical stability, we modify the erosion law \eqref{eqn:erosionModel} to include a term that penalizes curvature $\kappa$. The idea is to prevent regions of extremely high curvature from developing, as these regions would compromise the accuracy and stability of our fluid solver. We thus modify \eqref{eqn:erosionModel} follows
\begin{align}
\label{eqn:Vn}
\Vn = \atau + \eps \mean{\atau} \left( \frac{L }{2\pi} \kappa - 1 \right).
\end{align}
The parameter $\eps \ll 1$ controls the strength of the curvature-penalization. This term causes regions of high curvature to recede more rapidly and thus suppresses any high-frequency oscillations that may exist. Above, $\mean{\cdot}$ indicates a boundary mean. Since a closed body has mean curvature $\mean{\kappa} = 2\pi / L$, the penalty term has mean zero and and thus preserves area. The only source of material loss is therefore the shear-stress itself. In addition, we have scaled the penalty term on the mean absolute shear, $\mean{\atau}$, so that the ratio of smoothing-to-erosion always remains fixed, even as bodies vanish.

@@

The boundary of each body will be represented by a grid of collocation points. If we were to simply evolve the interfaces under \eqref{eqn:Vn2}, this grid would become distorted, which, if unchecked, could lead to severe tangling and numerical instability. We therefore introduce an artificial tangential velocity $\Vs$ that always keeps the grid points equally spaced in arclength \cite{hou-low-she1994}. The appropriate tangential velocity satisfies 
\begin{align}
\pderiv{\Vs}{\alpha} = \thalpha \Vn - \mean{\thalpha \Vn} \, ,
\end{align}
The evolution of the interface is then given by the combination of the normal and tangential velocity
\begin{align}
\dot{\xx}(t) = \Vn \nn + \Vs \ss \, .
\end{align}
Since $\Vs$ is tangential, it does not modify the physics of erosion, only the parameterization of the boundaries.

@@

With the shear stress computed, we are now ready to numerically evolve the boundaries using the {\thL} framework. The main equations to be solved are \eqref{eqn:Levo1}--\eqref{eq:xsurf}.  In Section~\ref{sec:thetaL} we introduced a curvature-penalizing term into the erosion law \eqref{eqn:Vn}. This term suppresses high-frequency oscillations, but also introduces {\em stiffness} into evolution equation \eqref{eqn:thetaevo1} for $\theta(\alpha,t)$. For numerical stability, we will treat the stiff term with an exponential integrator, while handling all other terms in \eqref{eqn:Levo1}--\eqref{eq:xsurf} with an explicit Runge-Kutta method. Ultimately, we will arrive at a method that is second-order in time.
